Influence bias in cultural evolutionary dynamics

### 2023-11-06 ms/ms.tex
some revisions with SE

Henrich and Broesch \citep{fijian_social_bias}  studied such indirect success biases, which they call cross-domain success bias or \emph{prestige bias} (e.g., great fishermen may be chosen as role-models for growing yams). They {suggested} that such biases, over generations, can lead to cultural adaptations, and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits
\citep[Ch.~8]{evolutionBook}. {Prestige bias may be} more common in humans than success {bias} \citep{complexityPaper}, and prestige is often mentioned in the cultural evolution literature, {although} there are few models of {it}.

@@

{Jimenez \& Mesoudi} (2019) \citep{prestigeSocialLearn} suggest that cultural evolution theory is a major factor in human ecological success. They mention that a way to acquire adaptive social information is by preferentially copying competent individuals within a valuable domain, and they call it success bias as we do in this paper. They also mention that competence within a domain is often difficult or impossible to directly asses, therefore people tend to use indirect cues of success, as suggested by {Henrich and Gil-White}.

In our paper, we examine the effect of both success and prestige biases, and their relationship in contribution to the cultural evolution of populations. We call the joint contribution of both biases \emph{Influence bias}.
This is an important distinction, as prestige is a context-dependent bias, rather than a content-{dependent} bias: it does not depend on the phenotype itself but rather on the number of copiers that have already copied each role-model, which may be easier and more accurate to estimate than success. 
Prestige bias is also frequency independent (see \Cref{cor:GBD} below), and thus it differs from conformity, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Optimal success-bias weight}

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal success-bias weight, $\alpha^*$, which minimizes the expected squared error between the chosen trait value and the optimal trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^* = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose role-model~$j$.
Indeed, if the trait value would have been correlated with fitness, the optimal success-bias weight would increase fitness. However in here we don't include natural selection.

@@

>>>>

Henrich and Broesch \citep{fijian_social_bias} studied such indirect success biases, which they call cross-domain success bias or \emph{prestige bias} (e.g., great fishermen may be chosen as role-models for growing yams). They suggested that such biases, over generations, can lead to cultural adaptations, and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits
\citep[Ch.~8]{evolutionBook}.
Prestige bias may be more common in humans than success bias \citep{complexityPaper}, and prestige is often mentioned in the cultural evolution literature, although there are few models of it; see Chellappoo (2021) for a philosophical critique of the concept of prestige itself. % TODO convert to reference; add this sentence to response letter.

@@

Jimenez \& Mesoudi~\citep{prestigeSocialLearn} suggest that cultural evolution theory is a major factor in human ecological success. They mention that a way to acquire adaptive social information is by preferentially copying competent individuals within a valuable domain, and they call it success bias as we do in this paper. They also mention that competence within a domain is often difficult or impossible to directly asses, therefore people tend to use indirect cues of success, as suggested by Henrich and Gil-White~\citep{prestige_evolution}.

Henrich and Gil-White~\citep{prestige_evolution} noted that ``the most skilled/knowledgeable role-models will, on average, end up with the biggest and most lavish clienteles, so the size and lavishness of a given model's clientele size (the \emph{prestige}) provides a convenient and reliable proxy for that person's information quality''. Therefore, they predicted that skilled individuals have higher status, people preferentially copy them more than others, and therefore prestigious individuals may be influential even beyond their domain of expertise.

Here, we examine the effect of both success and prestige biases, and their relationship in contribution to the cultural evolution of populations. We call the joint contribution of both biases \emph{influence bias}.
This is an important distinction, as prestige is a context-dependent bias, rather than a content-dependent bias: it does not depend on the phenotype itself but rather on the number of copiers that have already copied each role-model, which may be easier and more accurate to estimate than success. 
Prestige bias is also frequency independent (see \Cref{cor:GBD} below), and thus it differs from conformity, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Optimal success-bias weight}
% TODO edit the next sentence.
We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal success-bias weight, $\alpha^*$, which minimizes the expected squared error between the chosen trait value and the optimal trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^* = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose role-model~$j$.
Simply put, each copier estimates what success-bias weight $\alpha^*$ will result in copying a trait that is most similar to the ideal trait value $\hat{A}$.
Indeed, if the trait value is correlated with fitness, the optimal success-bias weight would increase the fitness of individuals. However, here we ignore the effects of natural selection, focusing instead on selectively neutral traits.

@@

Our decision to research and analyze a multi-bias model is preceded by several others.
Ze Hong~\citep{payoff}, for example, presented a model that combines both frequency-dependent bias (including conformist bias) and payoff bias. He has shown that an integrated learning strategy may be superior to relying on either bias alone. Moreover, both are likely to contribute to the spread of variants with high payoff, and the magnitude of conformist bias affects the evolutionary dynamics.
% TODO add Denton; others
